New method to describe the zonal symmetries and asymmetries of the Southern Annular Mode

### 2020-09-23 analysis/paper/paper.tex
casi casi

\title{Title here}

@@

\affiliation{CIMA UBA blablabla}

@@

yada yada SAM yada yada circulation.. yada yada so important. yada yada
many impacts.

\citet{fogt2012} studied the characteristics of the asymmetric structure
of the SAM. It computed the zonally anomalous component of mean sea
level pressure (MSLP) composites for positive and negative SAM events
and created two indices by projecting MSLP fields ontno them. However,
the use of composites leads to some issues that affects the
interpretability of the results. First, they can be dependent on the
choice of threshold used to define positive and negative events.
Secondly, by discarting data that don't meet the threshold, they don't
use all the information avaiable. Due to the realtive short timeframe
used, this leads to some composites being composed of as little as 4
years. Third, the resulting composites corresponding to each polarity
and season are derived from the average of different amount of fields
and from different years. This last issue is particularly important in
light of the changing structure of the SAM before and after 1980
\citep{silvestri2009}. In \citet{fogt2012}, the DJF SAM+ composite uses
only 7 years, 5 of which are later than 1988, whereas all of the 8 years
used for their DJF SAM- composite are from earlier than 1988.

@@

To separate between the zonally symmetric and asymmetric components of
the SAM, we computed the zonal mean and anomalies of the full SAM
spatial pattern, as shown in Figure~\ref{fig:method} for 700hPa. The
full spatial signal (\(\mathrm{EOF_1}(\lambda, \phi)\)) is the sum of
the zonally asymmetric (\(\mathrm{EOF_1^*}(\lambda, \phi)\)) and
symmetric (\([\mathrm{EOF_1}](\lambda, \phi)\)) components. We then
compute the ``Full'', ``Asymmetric'' and ``Symmetric'' indices, by
regressing each geopotential field on the respective patterns (weighting
by the cosine of latitude). The three indices are normalised by dividing
them by the standard deviation of the ``Full'' index at each level. As a
result, the magnitude between indices is comparable. However, only
``Full'' index will have unit standard deviation per definition.

\begin{figure*}
\includegraphics{method-1} \caption[Spatial patterns of the first EOF of 700 hPa geopotential height]{Spatial patterns of the first EOF of 700 hPa geopotential height. Full field (left), zonally asymmetric component (middle) and zonally symmetric component (right). Arbitrary units.}\label{fig:method}
\end{figure*}

@@

\begin{figure*}
\includegraphics{asymsam-timeseries-1} \caption[Time series for the asymmetric SAM and symmetric SAM and density estimates]{Time series for the asymmetric SAM and symmetric SAM and density estimates.}\label{fig:asymsam-timeseries}
\end{figure*}

@@

From Figure \ref{fig:asymsam-timeseries} we can see that the Asymemtric
and Symmetric time series appear to be correlated. Moreover, looking at
the extremes in the stratosphere, the Symmetric serie appears to lag the
Asymmetric series (see, for example, the positive events on late 1987).
We show these correlations, across all the levels of the reanalysis for
zero and -1 lag (Asymmetric index leading the Symmetric index), in
Figure \ref{fig:cor-lev}. Zero-lag correlations between the Asymmetric
and Symmetric series are relatively constant throught the troposphere,
fluctiating between 0.39 and 0.45. One-month-lag correlations are
similarly constant but significantly reduced to around 0.17. In the
stratosphere, zero-lag correlations drop to a minimum of 0.21 at 20 hPa
and then it increases again monotonically with height up to the
uppermost level of the reanalysis (although results near the top of the
models are to be interpreted with care). At the same time, one-month-lag
correlations increase with height. As a consequence, statospheric
Asymmetric index tend to precede corresponding Symmetric index.

\begin{figure}
\includegraphics{cor-lev-1} \caption[Correlation between the Symmetric and Asymmetric SAM at each level for lag zero and lag -1 (Asymmetric leads Symmetric)]{Correlation between the Symmetric and Asymmetric SAM at each level for lag zero and lag -1 (Asymmetric leads Symmetric).}\label{fig:cor-lev}
\end{figure}

\begin{figure*}
\includegraphics{cross-correlation-1} \caption[Cross correlation between levels of the Full, Asymmetric and Symmetric SAM]{Cross correlation between levels of the Full, Asymmetric and Symmetric SAM.}\label{fig:cross-correlation}
\end{figure*}

@@

\begin{figure*}
\includegraphics{trends-1} \caption[Decadal normalised trends for each index at each level for annual (row a) and seasonal values (rows b-e) for the period 1979-2018]{Decadal normalised trends for each index at each level for annual (row a) and seasonal values (rows b-e) for the period 1979-2018. Shading indicates the 95\% confidence interval.}\label{fig:trends}
\end{figure*}

@@

\begin{figure}
\includegraphics{r-squared-trend-1} \caption[Decadal trends for explained variance of each index at each level for the period 1979-2018]{Decadal trends for explained variance of each index at each level for the period 1979-2018. Shading indicates the 95\% confidence interval.}\label{fig:r-squared-trend}
\end{figure}

@@

\begin{figure*}
\includegraphics{2d-regr-1} \caption[Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM]{Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef. Points marked on panel b.2 are the location of the reference points used by \cite{raphael2004} for their Zonal Wave 3 index.}\label{fig:2d-regr}
\end{figure*}

@@

\begin{figure}
\includegraphics{wave-amplitude-1} \caption[Planteray wave amplitude for the regression patterns at 50 and 700 hPa]{Planteray wave amplitude for the regression patterns at 50 and 700 hPa. Note the varying x axis.}\label{fig:wave-amplitude}
\end{figure}

@@

\begin{figure}
\includegraphics{vertical-regression-1} \caption[Asymmetric coefficient of the multiple regression of mean monthly geopotential height anomalies between 65 and 40 South]{Asymmetric coefficient of the multiple regression of mean monthly geopotential height anomalies between 65 and 40 South. (\#FIXME this caption needs some love)}\label{fig:vertical-regression}
\end{figure}

@@

The wave-3 pattern from Figure \ref{fig:2d-regr} panel b.2 is very
similar to the Pacific-South American Pattern \citep{mo1987, kidson1988}
which is a teleconnection pattern associated with the ENSO
\citep{karoly1989}. Indeed, \citet{fogt2011} showed that there is a
significant relationship between the SAM and the ENSO. The correlation
between the full SAM and the ENSO as measured by the Oceanic Niño Index
\citep{bamston1997} (ONI) is -0.16. This relationship is captured
entirely the Asymmetric SAM, as this index has a partial correlation of
-0.26 with the ONI controlling for the effect of the Symmetric SAM,
whereas the Symmetric SAM's partial correlation with the ONI is
essentially null (0.019). We performed the same analysis using the
Multivariate Enso Index \citep{wolter2011} and the Southern Oscillation
Index \citep{ropelewski1987} to conclude that these results do not
depend on the ENSO index used.

@@

\begin{figure*}
\includegraphics{regr-air-season-1} \caption[Regression pattern of surface temperature with Asymmetric and Symmetric SAM]{Regression pattern of surface temperature with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas. Gray areas have more than 15\% of missing data.}\label{fig:regr-air-season}
\end{figure*}

@@

\begin{figure*}
\includegraphics{pp-regr-oceania-1} \caption[Regression pattern of precipiation with Asymmetric and Symmetric SAM]{Regression pattern of precipiation with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas.}\label{fig:pp-regr-oceania}
\end{figure*}

@@

\begin{figure*}
\includegraphics{pp-regr-america-1} \caption[Same but for america]{Same but for america}\label{fig:pp-regr-america}
\end{figure*}

@@

Temperature anomalies associated with the Full SAM broadly show a
pattern of negative anomalies at polar latitudes surrounded by positive
anomalies, but with many deviations from symmetry. The Asymemtric SAM
index explains a big portion of these deviations. In particular, the
positive phase of the Asymmetric SAM is associated with colder
temperatures over Southern Brazil, South Africa and Southern Australia,
as well as the negative anomalies in the equatorial Pacific consistent
with the ENSO-SAM relationship delineated above. These are particularly
clear in the DJF and SON trimesters, which include the months in which
the ENSO teleconnection is more active \citep[\citet{fogt2011},
\citet{cai2020a}]{cazes-boezio2003}.

@@

\begin{figure}
\includegraphics{A1-1} \appendcaption{A1}{Lag-correlation between Symmetric and Asymmetric SAM at each level.}\label{fig:A1}
\end{figure}

\begin{figure}
\includegraphics{A2 ccf-levels-1} \caption[Cross-correlation functions for each index and two differnet base levels]{Cross-correlation functions for each index and two differnet base levels.}\label{fig:A2 ccf-levels}
\end{figure}

\begin{figure}
\includegraphics{A3-1} \appendcaption{A3}{Fourier spectrum of each timeseries. The shading indicates de 95\% area derived by fitting an AR process to each series and bootstrapping 5000 simulated samples.}\label{fig:A3}
\end{figure}

\begin{figure}
\includegraphics{A8-1} \appendcaption{A8}{Pattern cross-correlation \#FIXME!}\label{fig:A8}
\end{figure}

\begin{figure}
\includegraphics{A9-1} \caption[700 hPa Geopotetnial height zonal anomalies of composites of positive and negative SAM months selected using 1 standard deviation as threshhold]{700 hPa Geopotetnial height zonal anomalies of composites of positive and negative SAM months selected using 1 standard deviation as threshhold. Numbers in the column headers are pattern correlation between SAM+ and SAM- composites and number of monthly fields used to construct the composites.}\label{fig:A9}
\end{figure}

\begin{verbatim}
#> # A tibble: 12 x 3
#>    item1 item2 correlation
#>    <fct> <fct>       <dbl>
#>  1 MAM   DJF     -1.16e-33
#>  2 JJA   DJF      3.07e-34
#>  3 SON   DJF      2.55e-33
#>  4 DJF   MAM     -1.16e-33
#>  5 JJA   MAM      1.84e-33
#>  6 SON   MAM      1.53e-32
#>  7 DJF   JJA      3.07e-34
#>  8 MAM   JJA      1.84e-33
#>  9 SON   JJA     -4.03e-33
#> 10 DJF   SON      2.55e-33
#> 11 MAM   SON      1.53e-32
#> 12 JJA   SON     -4.03e-33
\end{verbatim}

\begin{figure}
\includegraphics{A10-1} \caption[Zonal of projection of 700 hPa onto the first EOF of each season]{Zonal of projection of 700 hPa onto the first EOF of each season.}\label{fig:A10}
\end{figure}

\begin{figure}
\includegraphics{A11-1} \caption[Spatial patterns of the first EOF of 700 hPa geopotential height computed for the periods 1979 to 1998 and 1999 to 2018]{Spatial patterns of the first EOF of 700 hPa geopotential height computed for the periods 1979 to 1998 and 1999 to 2018. Full field (left), zonally asymmetric component (middle) and zonally symmetric component (right). Pattern correlation between both fields is 0.77. Arbitrary units.}\label{fig:A11}
\end{figure}

>>>>

\title{New method to describe the zonal symmetries and asymmetries of the
Southern Annular Mode}

@@

\affiliation{Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales,
Departamento de Ciencias de la Atmósfera y los Océanos, Buenos Aires,
Argentina CONICET -- Universidad de Buenos Aires, Centro de
Investigaciones del Mar y la Atmósfera (CIMA), Buenos Aires, Argentina
CNRS -- IRD -- CONICET -- UBA, Instituto Franco‐Argentino para el
Estudio del Clima y sus Impactos (UMI 3351 IFAECI), Buenos Aires,
Argentina}

@@

The Southern Annular Mode (SAM) is the main mode of variability in the
Southern Hemisphere extratropical circulation \citep{rogers1982} in
daily, monthly, and decadal timescales {[}\citet{baldwin2001a};
fogt2006{]} and exerts an important influence in weather conditions such
as temperature and precipitation anomalies and sea ice concentration
\citep{fogt2020}. Its positive phase is traditionally described as
anomalously low pressures over Antarctica surrounded by a ring of
anomalous high pressures in middle-to-high latitudes.

However, computed as the leading Empirical Orthogonal Function (EOF) of
Sea Level Pressure or low-level geopotential height, the SAM spatial
structure contains noticeable deviations from this zonally symmetric
description, particularly in the Pacific Ocean region. These zonal
asymmetries are not widely studied, but previous work suggest that they
strongly modulate the regional impacts of the SAM, going as far as
reversing its relationship between precipitation in South America
\citep{silvestri2009}. At the very least, the fact that the SAM is not
entirely zonally symmetric hinders our ability to reconstruct its
historical variability prior to the availabilty of dense observations in
the Southern Hemisphere \citep{jones2009}.

We are not aware of any previous work which quantifies the temporal
variability of the asymmetric component of the SAM with the exception of
\citet{fogt2012}. However, their methods based on composites of positive
and negative SAM events leads to some issues, such as spatial patterns
derived from as little as 4 cases and from imbalanced periods (for
example, 5 of the 7 cases in their DJF SAM+ composite are from later
than 1988, whereas all of the 8 years in their DJF SAM- composite are
from earlier than 1988). This is particularly important due to the
inhomogeneities in renalysis products prior to the sattelite era and the
possible change in the asymmetric structure of the SAM
\citep{silvestri2009}.

Our objective is, then, to systematically characterise the zonally
asymmetric component of the SAM variability by constructing two indices
which aim to capture exclusively the variability of the symmetric and
asymmetric component each. We then analyse their temporal variability,
trends and vertical coherence. We study the spatial patterns described
by the variability exclusive to each index. Finally, we investigate
theri relationship with temperature and recipitation anomalies.

@@

To separate between the zonally symmetric and asymmetric components of
the SAM, we computed the zonal mean and anomalies of the full SAM
spatial pattern, as shown in Figure~\ref{fig:method} for 700hPa. The
full spatial signal (\(\mathrm{EOF_1}(\lambda, \phi)\)) is the sum of
the zonally asymmetric (\(\mathrm{EOF_1^*}(\lambda, \phi)\)) and
symmetric (\([\mathrm{EOF_1}](\lambda, \phi)\)) components. We then
compute the ``Full SAM'', ``Asymmetric SAM'' and ``Symmetric SAM''
indices as the regression coefficients of the regression ofeach monthly
geopotential field on the respective patterns (weighting by the cosine
of latitude). The three indices are normalised by dividing them by the
standard deviation of the ``Full'' index at each level. As a result, the
magnitude between indices is comparable. However, only ``Full'' index
will have unit standard deviation per definition. From the regression,
we also use the explained variance of each pattern as a indication of
the degere of symmetry or asymmetry of each monthly field.

\begin{figure*}
\includegraphics{method-1} \caption[Spatial patterns of the first EOF of 700 hPa geopotential height]{Spatial patterns of the first EOF of 700 hPa geopotential height. (a) Full field, (b) zonally asymmetric component and (c) zonally symmetric component. Arbitrary units.}\label{fig:method}
\end{figure*}

@@

\begin{figure*}
\includegraphics{asymsam-timeseries-1} \caption[Time series for the Asymmetric SAM and Symmetric SAM indices at (a) 50 hPa and (b) 700 hPa]{Time series for the Asymmetric SAM and Symmetric SAM indices at (a) 50 hPa and (b) 700 hPa. To the left, probability density estimate of each index. Series are standarised by the standard deviation of the Full SAM at each level.}\label{fig:asymsam-timeseries}
\end{figure*}

@@

From Figure \ref{fig:asymsam-timeseries} we can see that the Asymemtric
and Symmetric time series appear to be correlated. Moreover, looking at
the extremes in the stratosphere, the Symmetric serie appears to lag the
Asymmetric series (see, for example, the positive events on late 1987).
We show these correlations, across all the levels of the reanalysis for
zero and -1 lag (Asymmetric index leading the Symmetric index), in
Figure \ref{fig:cor-lev}. Zero-lag correlations between the Asymmetric
and Symmetric series are relatively constant throught the troposphere,
fluctiating between 0.39 and 0.45. One-month-lag correlations are
similarly constant but significantly reduced to around 0.17. In the
stratosphere, zero-lag correlations drop to a minimum of 0.21 at 20 hPa
and then it increases again monotonically with height up to the
uppermost level of the reanalysis (although results near the top of the
models are to be interpreted with care). At the same time, one-month-lag
correlations increase with height. As a consequence, statospheric
Symmetric index tend to precede corresponding Asymmetric index.

\begin{figure}
\includegraphics{cor-lev-1} \caption[Correlation between the Symmetric SAM and Asymmetric SAM index at each level for lag zero and lag -1 (Symmetric leads Asymmetric)]{Correlation between the Symmetric SAM and Asymmetric SAM index at each level for lag zero and lag -1 (Symmetric leads Asymmetric).}\label{fig:cor-lev}
\end{figure}

\begin{figure*}
\includegraphics{cross-correlation-1} \caption[Cross correlation between levels of the (a) Full SAM, (b) Asymmetric SAM, and (c) Symmetric SAM]{Cross correlation between levels of the (a) Full SAM, (b) Asymmetric SAM, and (c) Symmetric SAM.}\label{fig:cross-correlation}
\end{figure*}

@@

\begin{figure*}
\includegraphics{trends-1} \caption[Decadal trends at each level for annual (row 1) and seasonal values (rows 2 to 5) for the period 1979-2018 and for the (column a) Full SAM index, (column b) Asymmetric SAM index, and (column c) Symmetric SAM index]{Decadal trends at each level for annual (row 1) and seasonal values (rows 2 to 5) for the period 1979-2018 and for the (column a) Full SAM index, (column b) Asymmetric SAM index, and (column c) Symmetric SAM index. Shading indicates the 95\% confidence interval.}\label{fig:trends}
\end{figure*}

@@

\begin{figure}
\includegraphics{r-squared-trend-1} \caption[Decadal trends of the variance explained by the Asymmetric and Symmetric SAM at each level for the period 1979-2018]{Decadal trends of the variance explained by the Asymmetric and Symmetric SAM at each level for the period 1979-2018. Shading indicates the 95\% confidence interval.}\label{fig:r-squared-trend}
\end{figure}

@@

\begin{figure*}
\includegraphics{2d-regr-1} \caption[Regression patterns of geopotential height (meters) at (row 1) 50 hPa and (row 2) 700 hPa with the (column a) Full SAM, (column b) Asymmetric SAM, and (column c) Symmetric SAM]{Regression patterns of geopotential height (meters) at (row 1) 50 hPa and (row 2) 700 hPa with the (column a) Full SAM, (column b) Asymmetric SAM, and (column c) Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef. Points marked on panel b.2 are the location of the reference points used by \cite{raphael2004} for their Zonal Wave 3 index. }\label{fig:2d-regr}
\end{figure*}

@@

\begin{figure}
\includegraphics{wave-amplitude-1} \caption{Amplitude (meters) of zonal waves of the geopotential height regression patterns in Figure \ref{fig:2d-regr} for zonal waves with wave-number 0, 1, 2, and 3, where wave-number 0 represents the amplitude of the zonal mean. Note the differnet x axis.}\label{fig:wave-amplitude}
\end{figure}

@@

\begin{figure}
\includegraphics{vertical-regression-1} \caption[Regression between monthly geopotential anomalies (meters) averaged betweeen 65\degree and 40\degree S and the Asymmetric SAM index (extracted from multiple regression including the Symmetric SAM)]{Regression between monthly geopotential anomalies (meters) averaged betweeen 65\degree and 40\degree S and the Asymmetric SAM index (extracted from multiple regression including the Symmetric SAM). (a) With the Asymmetric SAM in 50 hPa and (b) in 700 hPa.}\label{fig:vertical-regression}
\end{figure}

@@

The wave-3 pattern from Figure \ref{fig:2d-regr} panel b.2 is very
similar to the Pacific-South American Pattern \citep{mo1987, kidson1988}
which is a teleconnection pattern associated with the ENSO
\citep{karoly1989}. Indeed, \citet{fogt2011} showed that there is a
significant relationship between the SAM and the ENSO. The correlation
between the full SAM and the ENSO as measured by the Oceanic Niño Index
\citep{bamston1997} (ONI) is -0.16. Consistent with \citet{fan2007}, we
show that this relationship is captured entirely the Asymmetric SAM, as
this index has a partial correlation of -0.26 with the ONI controlling
for the effect of the Symmetric SAM, whereas the Symmetric SAM's partial
correlation with the ONI is essentially null (0.019). We performed the
same analysis using the Multivariate Enso Index \citep{wolter2011} and
the Southern Oscillation Index \citep{ropelewski1987} to conclude that
these results do not depend on the ENSO index used.

@@

\begin{figure*}
\includegraphics{regr-air-season-1} \caption[Regression pattern of season mean surface temperature (Kelvin) with Asymmetric SAM and Symmetric SAM]{Regression pattern of season mean surface temperature (Kelvin) with Asymmetric SAM and Symmetric SAM. Black contours indicate areas with p-value smaller than 0.05 controlling for False Detection Rate. Gray areas in Antarctica are areas with have more than 15\% of missing data.}\label{fig:regr-air-season}
\end{figure*}

@@

\begin{figure*}
\includegraphics{pp-regr-oceania-1} \caption[Regression pattern of (row 1) annual and (rows 2 to 5) season mean precipitation anomalies (mm per day, shading) and 700 hPa geopotential height (thin lines, positive values as solid lines and negative values as dashed lines) with (column a) Full SAM, (column (b) Asymmetric SAM and (column c) Symmetric SAM]{Regression pattern of (row 1) annual and (rows 2 to 5) season mean precipitation anomalies (mm per day, shading) and 700 hPa geopotential height (thin lines, positive values as solid lines and negative values as dashed lines) with (column a) Full SAM, (column (b) Asymmetric SAM and (column c) Symmetric SAM. thin lines are the Black contours indicate areas with p-value smaller than 0.05 controlling for False Detection Rate.}\label{fig:pp-regr-oceania}
\end{figure*}

@@

\begin{figure*}
\includegraphics{pp-regr-america-1} \caption{Same as Figure \ref{fig:pp-regr-oceania} but for South America.}\label{fig:pp-regr-america}
\end{figure*}

@@

Temperature anomalies associated with the Full SAM broadly show a
pattern of negative anomalies at polar latitudes surrounded by positive
anomalies, but with many deviations from symmetry. The Asymemtric SAM
index explains a big portion of these deviations. In particular, the
positive phase of the Asymmetric SAM is associated with colder
temperatures over Southern Brazil, South Africa and Southern Australia,
as well as the negative anomalies in the equatorial Pacific consistent
with the ENSO-SAM relationship delineated above. These are particularly
clear in the DJF and SON trimesters, which include the months in which
the ENSO teleconnection is more active
\citep{cazes-boezio2003, fogt2011, cai2020a}.

@@

\begin{figure}
\includegraphics{A1-1} \appendcaption{A1}{Lag-correlation between Asymmetric SAM and Symmetric SAM index at each level. Negative lags imply Symmetric SAM leading Asymmetric SAM and vice versa.}\label{fig:A1}
\end{figure}

\begin{figure}
\includegraphics{A3-1} \appendcaption{A3}{Fourier spectrum of each timeseries computed as Fourier transform smoothed with modified Daniell smoothers with withs 3 and 5. The shading indicates de 95\% confidence area derived by fitting an autorregressive model and computing the spectrum for 5000 simulated samples from the fitted autoregressive model (95\% of the simulated sampels had an amplitude equal or lower). The light line indicates the theoretical expected amplitude from the autorregressive model.}\label{fig:A3}
\end{figure}

\begin{figure}
\includegraphics{A9-1} \caption[700 hPa Geopotetnial height zonal anomalies (meters) of composites of positive and negative SAM months selected using $\pm1$ standard deviation as threshhold]{700 hPa Geopotetnial height zonal anomalies (meters) of composites of positive and negative SAM months selected using $\pm1$ standard deviation as threshhold. Numbers in the column headers are pattern correlation between SAM+ and SAM- composites and number of monthly fields used to construct the composites.}\label{fig:A9}
\end{figure}

\begin{figure}
\includegraphics{A10-1} \caption[Regression of 700 hPa geopotential height zonal anomalies (meters) onto the standarised timeseries of the leading EOF computed for each season independently]{Regression of 700 hPa geopotential height zonal anomalies (meters) onto the standarised timeseries of the leading EOF computed for each season independently.}\label{fig:A10}
\end{figure}

\begin{figure}
\includegraphics{A11-1} \caption[Regression of 700 hPa geopotential height zonal anomalies (meters) onto the standarised timeseries of the leading EOF computed for the periods 1979 to 1998 and 1999 to 2018]{Regression of 700 hPa geopotential height zonal anomalies (meters) onto the standarised timeseries of the leading EOF computed for the periods 1979 to 1998 and 1999 to 2018. Pattern correlation between both fields is 0.76.}\label{fig:A11}
\end{figure}
